---
title: "Project Proposal 2"
author:  
  - Jake Kaihewalu (4012753)
  - Jasmine Kellogg (4266367)
  - Andrew Nguyen (5495395)
  - Masaki Osato (4222154)
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output: pdf_document
---
# Research Question 

Can we classify whether or not a patient has cardiovascular disease based on a variety of catergorical and numeric variables?

# Data

Our response variable will be the binary classification variable "cardio", which is 1 if the patient tests positive and 0 if they test negative. There are 11 predictor variables, 6 of which are categorical and 5 are numeric. 

# Analysis Plan

Firstly, we plan on using exploratory data analysis and graphics to find whether some predictors are highly correlated to one another or insignificant in its effect on the response variable. We then plan on using logistic regression, classification trees and random forests to classify observations into two binary groups, splitting based on largest reductions in impurity (using either the Gini index or entropy). For the logistic regression, we can fit a model based on a new variable (cardio) that indicates whether or not a patient tested positive for cardiovascular disease. We plan to use 70% of our dataset for training and 30% of our dataset for testing since we have a significant amount of observations. 

# References 

Cardiovascular Disease Dataset. Ulianova, Svetlana. Kaggle, 20 Jan. 2019, [www.kaggle.com/sulianova/cardiovascular-disease-dataset.]

\newpage

# Data Overview 

```{r}
#Read in data
library(readr)
cardio_train <- read.csv("~/131-final-project/Final Project/cardio_train.csv", sep=";")
#View(cardio_train)
```


```{r}
## Split the data 
set.seed(3)
id <- sample(1:nrow(cardio_train), 0.75*nrow(cardio_train))
cardio_train.train <- cardio_train[id,]
cardio_train.test <- cardio_train[-id,]
dim(cardio_train.train)
dim(cardio_train.test)
```

```{r}
## Summarize response
summary(cardio_train.train$cardio)
```

```{r}
# Basic model
attach(cardio_train)
glm.fit <- glm(cardio ~ age + gender + height + weight + ap_hi + ap_lo + cholesterol
               + gluc + smoke + alco + active)
summary(glm.fit)
```